Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-01 demystify.tex
updates

\begin{itemize}
\item \textbf{Too personal} Don't talk about your personal issues, e.g., family, health, relationship, etc. This can raise concerns about your ability to handle the PhD. Don't talk about religious or political beliefs (\autoref{sec:cultural}).
\item \textbf{Criticizing} your current or previous institution or professors: just like in a job interview, don't badmouth your current or previous employer. 
\item \textbf{No concrete evidence to back up claims}: e.g., saying you are passionate about a research topic without showing any experience. This is where specific names (work submitted at conf. X), numbers (outperformed SOTA by Y \%), and examples (worked on project Z) can help. 
\item \textbf{Use flowery and AI-like language}: don't use AI to write your SOP. Not that hard to raise suspicion.
\item \textbf{Not customized to the program}: If your SOP can be sent to multiple programs with few changes, it is too generic. Do some homework and mention why you want to spend the next 5--7 years there.
\item \textbf{Mentioning wrong professors}: do not mention emeritus professors or those who have left.  Do your homework to see if the professors you mention are still active in research.  Teaching and adjunct faculty are often not active in mentoring PhD students (\autoref{sec:faculty-types}).
\item \textbf{Too Long and fancy format}: As mentioned, keep it under 2 pages. Don't use too much coloring or fancy fonts (like those in Words). CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX. 
\end{itemize}

>>>>

\begin{itemize}
\item \textbf{Too personal} Don't talk about your personal issues, e.g., family, health, relationship, etc. This can raise concerns about your ability to handle the PhD. Don't talk about religious or political beliefs (\autoref{sec:cultural}).
\item \textbf{Criticizing} your current or previous institution or professors. Just like in a job interview, don't badmouth your current or previous employer. 
\item \textbf{No concrete evidence to back up claims}: e.g., saying you are passionate about a research topic without showing any experience. This is where specific names (work submitted at conf. X), numbers (outperformed SOTA by Y \%), and examples (worked on project Z) can help. 
\item \textbf{Use flowery and AI-like language}. Don't use AI to write your SOP. Not that hard to raise suspicion. Though you can ask AI to check your grammar and spelling.
\item \textbf{Not customized to the program}: If your SOP can be sent to multiple programs with few changes, it is too generic. Do some research and mention why you want to spend the next 5--7 years there.
\item \textbf{Mentioning wrong professors}: do not mention emeritus professors or those who have left.  Teaching and adjunct faculty are often not active in mentoring PhD students (\autoref{sec:faculty-types}). Do your homework and mention profs who are till active in research. 
\item \textbf{Too Long and fancy format}: As mentioned, keep it under 2 pages. Don't use too much coloring or fancy fonts (like those in Words). CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX (with Times or default font, 11pt, and 1-inch margin).
\end{itemize}
